Investigating the influence of genre preferences on movie/series rating patterns

### 2024-03-22 gen/paper/paper.Rmd
plots and pdfs

The entertainment industry is a dynamic industry. Consumer preferences
continuously change, and new movies and series are continuously being
produced. Genre continues to be an important topic in the entertainment
industry. Genre preferences indicate what consumers liked to watch in
the past, and therefore probably also would like to watch in the future.
Thus, genre can serve as a selection criteria, based on which consumers
decide whether to watch a certain movie or not. Therefore, an
understanding of genre preferences gives movie producers a direction on
what type of genre they should adopt to increase chances of success and
to best deliver the message they want to bring across. Consequently, it
is important to investigate the evolution of consumers’ genre
preferences over time.

@@

**How do consumers' movie genre preferences change over time?**

@@

In this paper we investigate consumers’ movie and series preferences
using five of IMDb's datasets containing movie data. First, these
datasets are loaded, cleaned, inspected, merged and cleaned. Next, the
data is analyzed using a regression analysis. This regression analysis
regresses all movie and serie genre's against the average rating's over
a period of 5 years. Additionally, the variables title, movieType and
time are used as control variables to improve the interpret ability of
the estimates.

| Name of Variable | Contents of Variable                                                                                                                                             | Used as .. in Analysis |
|------------------|------------------------------------------------------------------------------------------------------------------------------------------------------------------|------------------------|
| Average Rating   | The average rating on IMBd per title of a given movie or series.                                                                                                 | Dependent Variable     |
| Genre Types      | Genre Types is a cluster for all 26 genre dummies that contain the value 1 if the movie or series contains that genre and otherwise 0.                           | Independent Variable   |
| titleType        | This variable indicates whether a title is a movie or a series.                                                                                                  | Fixed Effect           |
| Title            | The title is the given title of any unique movie or series                                                                                                       | Fixed Effect           |
| 5 Year Window    | The 5 year windown (or ' slide_window_5' indicates between what 5 year window a movie or series belongs to. For example a movie from 2003 belonging to 2000-2004 | Split Varible          |

### Average Movie Rating

@@

### Genres

@@

## Results and Discussion
```{r}
#insert pdf of results graph with colored dots
```

@@

>>>>

The entertainment industry is highly dynamic. Consumer preferences keep
changing, and new movies and series are produced regularly. These
changes in preferences require the industry to respond accordingly. For
example, showing an ankle in a movie might have been considered
scandalous in the 1950s, but it is now just a minor detail. To make an
impact in such an ever-evolving industry, movie directors need to get
creative. These changes in the industry influence how movies and series
are produced and how consumers react to them. One way to understand
these patterns is by examining the genre of a product.

Genre preferences reflect what consumers have enjoyed watching in the
past, and hence are likely to enjoy in the future. Therefore, genre can
be used as a criterion for selecting productions, based on which
consumers decide whether to watch a particular production or not. An
understanding of genre preferences provides filmmakers with guidance on
which genre to choose to increase the chances of success and to best
convey their intended message. However, the nature of genres is
constantly evolving, and it is important to comprehend how current
consumers respond to the evolution of a genre over time. For instance,
an action movie today might have rapid scene changes, multiple plot
twists, and lengthy fight scenes, but on average, such movies may only
receive moderate ratings at best. In contrast, action movies from two
decades ago had a more gradual plot development and were less
overwhelming, yet they received a higher average rating. When consumers
rely on the genre productions of the past to make their choice, it may
adversely impact their evaluation of a current production in the same
genre.

In order to gain a better understanding of how modern consumers rate
products and services, it is crucial to comprehend how they use their
past preferences as a point of reference. Hence, this research paper
aims to explore whether there exist any discernible patterns in genre
preferences over time, and whether some genres tend to have a greater
impact on average ratings than others.

@@

**How do consumers' production preferences of genre elements change over
time, and how do they impact the average ratings of those same
productions?**

## Relevance

The research results shed light on how consumers use dynamic genres to
shape their satisfaction with a movie or TV series. The findings suggest
that some genres are preferred in certain years over others. For
example, war movies tend to receive higher ratings during world wars
compared to times of peace. This implies that consumers use these
productions as references to rate them in context of the sign of the
times. Managers may want to consider adjusting their productions to
better suit the elements in those preferred years.

The study also found that recent fantasy movies tend to receive lower
ratings from consumers, while those produced in the 70s are rated more
favorably. Overall, the results suggest that managers may face
challenges in meeting previously set expectations and that these
findings provide a reference frame for when those expectations are met.

Additionally, the results also provide insights into how genres are
related to average ratings across another. Currently, every production
has a similar pattern of ingredients. Most movies contain hints of drama
and romance. Whether it is in a distropian sci-fi survival movie or an
horror action series. However, the results indicated that these common
genres are not significantly related to the average ratings. In contrast
some movies and series do contain unique elements do results in
significant relationships. For example, biographies and animations. This
may imply that consumers may not rely so heavily on past genres that are
common because they gradually change with times without consumers
noticing. (Think of it like a birthday party with your family when you
are 12. Your aunt keeps telling you "How much you have grown!", but you
don't notice any difference) In contrast, the uniqueness of biographies
does sharply stand out. Therefore, managers may want to focus more
strongly on the less common genre features compared to more common ones.

@@

In this paper we investigate consumers’ movie and series preferences
using four of IMDb's datasets containing movie data. First, these
datasets are loaded, cleaned, inspected, merged and cleaned. Next, the
data is analyzed using a regression analysis. This regression analysis
regresses all movie and series genre's against the average rating's over
a sliding period of 5 years. Additionally, the variables title,
movieType and time are used as control variables to improve the
interpret ability of the estimates.

| Name of Variable | Contents of Variable                                                                                                                                             | Used as .. in Analysis |
|------------|-------------------------------------------------|------------|
| Average Rating   | The average rating on IMBd per title of a given movie or series.                                                                                                 | Dependent Variable     |
| Genre Types      | Genre Types is a cluster for all 26 genre dummies that contain the value 1 if the movie or series contains that genre and otherwise 0.                           | Independent Variable   |
| titleType        | This variable indicates whether a title is a movie or a series.                                                                                                  | Fixed Effect           |
| Title            | The title is the given title of any unique movie or series                                                                                                       | Fixed Effect           |
| 5 Year Window    | The 5 year windown (or ' slide_window_5' indicates between what 5 year window a movie or series belongs to. For example a movie from 2003 belonging to 2000-2004 | Split Varible          |

### Outcome Variable: Average Movie Rating

@@

### Explanatory Variable: Genres

@@

## Results and Discussion

```{r}
#insert pdf of results graph with colored dots
```

@@

```{r}
